---
author: "Jon Karanezi, Madeline Olson, Kaitlyn Schott, Ferris Wolf, Simon Wu" 
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(broman)
library(modelr)
library(kableExtra)
source("viridis.R")
source("ggprob.R")
```

```{r, include=FALSE}
amazon = read_csv("dataset.csv") %>%
    drop_na() %>%
    rename(
        amazon = position_first_amazon,
        partner = position_first_non_amazon,
        third_party = position_first_wholly_non_amazon,
        partner_stars = non_amazon_stars,
        partner_reviews = non_amazon_reviews,
        third_party_stars = wnon_amazon_stars,
        third_party_reviews = wnon_amazon_reviews
    )

amazon = amazon %>% 
  mutate(is_general = case_when(amazon <= 3 ~ "yes", amazon > 3 ~ "no"))
amazon

```

# What Characterizes a Product's Placement on Amazon?


## Introduction
Our dataset is a subset of search queries on Amazon from a GitHub repository created by The Markup on October 14, 2021.

The actual data retrieval was done on January 21st, 2021 from a single location, Washington D.C. 

Given the author's standards regarding Amazon brands, the data collected was narrowed to 3,492 different queries and accumulated different information based on each query. We are interested in this data because it relates Amazon products and their quality to those of Amazon partners and independent retailers.

Using this data, we came up with three research questions to help us explore the characteristics that define an Amazon search:

- Does what you search matter in terms of product ranking? 

- Can we assume the type of search by simply looking at the product's ranking, stars and reviews on Amazon? 

- How do Amazon's products compare to those marketed by non-Amazon retailers?

## Data
This data includes three main categories relating to the Amazon search queries—position on the results page, Amazon stars, and Amazon review count. 
For each search query, there is an associated amazon product, non_amazon product and independent product ranking, and associated star rating and review count for each product type. For ranking on Amazon, the highest possible place would be 1.0. As products approach 1.0, they are considered to be of more relevance to the search query. The stars are rated on a scale of 0 to 5, meaning as product score approaches 5, they are considered a better product. As the number of reviews increases, the item is reflected to increase in popularity and frequency. If there is a high number of reviews and a high star rating, that would indicate that the product is high-quality and tested out by many consumers, although there is a caveat to this mentioned in the ***limitations*** section

## Variables
```{r, echo = FALSE}

amazon_variables = tibble(
  Variables = c("search_term",
                "amazon",
                "partner",
                "third_party",
                "amazon_stars",
                "partner_stars",
                "third_party_stars",
                "amazon_reviews",
                "partner_reviews",
                "third_party_reviews"),
  Description = c("search term inputted into Amazon website",
                  "first Amazon product ranking associated with search",
                  "first parter product ranking associated with search",
                  "first third party product ranking associated with search",
                  "average star rating of respective Amazon product",
                  "average star rating of respective partner product",
                  "average star rating of respective third party product",
                  "number of reviews of respective Amazon product",
                  "number of reviews of respective partner product",
                  "number of reviews of respective third party product")
)
amazon_variables %>% 
  kable(caption = "Key Variables from the Amazon Search Term Dataset") %>% 
  kable_styling(position = "left", full_width = FALSE,
                bootstrap_options = c("striped"))
```

```{r, include = F}
# Summarizing Star and Reviews averages
summary = amazon %>% 
  select(-search_term) %>% 
  summarize(amazon = mean(amazon_stars),
            partner = mean(partner_stars),
            third_party = mean(third_party_stars))

summary1 = amazon %>% 
  select(-search_term) %>% 
  summarize(amazon = mean(amazon_reviews),
            partner = mean(partner_reviews),
            third_party = mean(third_party_reviews))

prop_third_party = amazon %>% 
  filter(third_party_stars < partner_stars & third_party_stars < amazon_stars)

prop_partner= amazon %>% 
  filter(partner_stars < third_party_stars & partner_stars < amazon_stars)

prop_amazon = amazon %>% 
  filter(amazon_stars < partner_stars & amazon_stars < third_party_stars)

tibble(
  prop_amazon = nrow(prop_amazon) / nrow(amazon),
  prop_third = nrow(prop_third_party) / nrow(amazon),
  prop_partner = nrow(prop_partner) / nrow(amazon)
)

# This snippet highlights how what you search affects the placement drastically

# If we consider products where Amazon does not have the 1st spot
study_amazon = amazon %>% 
  filter(amazon != 1) %>% 
  pull(amazon)

# The amazon products get placed very low
mean(study_amazon)
sd(study_amazon)


# In contrast, if we consider where amazon products are 1
study_non_amazon = amazon %>% 
  filter(amazon == 1) %>% 
  pull(partner)

# The partner products don't get placed as lower. Meaning, that even if the third party products aren't first, they are still placed closer to the top of results
mean(study_non_amazon)
sd(study_non_amazon)
```

### Finding a threshold
It would be a very time consuming task to try and predict the characteristics of a product through the search term, since there are limitless possibilities for what you could search. In our study, we instead looked at the data that each search term provided, to see whether we can characterize the product that way.

Upon inspection of the dataset, we notice that the vast majority of product names are split between two categories. More specific search terms like "13 inch cube storage bins" and more general search terms like "alcohol 70%". (how can u assume this when you don't have search volume data).

But how do we differentiate between these two?

We would need some threshold value for the positioning of products to decide what makes a product more general or specific.

The framework behind Amazon allows Amazon and other parties to sell products. We understand that Amazon is unable to create every possible product on earth because they are a single company. So, a majority of products are not Amazon, and it would make sense that for those products that are there, they would satisfy more general search terms, for a larger audience. There is always the case that a super specific search term has an associated Amazon product.

We will use the Amazon product positions to find a value suggesting that having a position at that value or lower, suggests that Amazon has its own product, which is why it is shown so high in the ranking.

```{r, echo = F}
threshold = amazon %>% 
  pull(amazon)

mu = mean(threshold)
sigma = sd(threshold)


pnorm(3, mean = mu, sd = sigma)
```

Considering the lower ~ 16% of the values, we come to a value of 3, which would say that a large majority of amazon products (84%) are positioned greater than 3. We end up using 3 as the threshold since it captures a realistic percentage of amazon products that are general, in our view. Any products placed with a value of 3 or lower are considered General Products, while all others (greater than 3) are considered Specific Products. 


### Question 1
- How do General and Specific Search Terms Position Averages Vary?

>We answer this by comparing general and specific search queries average position using boxplots between Amazon and partner products.

```{r, echo=FALSE}

#creates DF with only secific products
specific_search = amazon %>% 
  filter(is_general == "no") %>% 
  select(amazon, partner)

#pivot data to make box plots possible
specific_search = specific_search %>% 
  pivot_longer(cols=1:2, names_to = "product_type", values_to = "position")

ggplot(specific_search, aes(x=product_type, y=position)) +
  geom_boxplot(color = 'palegreen4', fill = 'hotpink1') +
  ggtitle("Position Average for Specific Products") +
  xlab("Product Type") +
  ylab("Position")

#creates DF with only general products
general_search = amazon %>% 
  filter(is_general == "yes") %>% 
  select(amazon, partner)

#pivot data to make box plots possible
general_search = general_search %>% 
  pivot_longer(cols=1:2, names_to = "product_type", values_to = "position")


ggplot(general_search, aes(x=product_type, y=position)) +
  geom_boxplot(color = 'palegreen4', fill = 'hotpink1') +
  ggtitle("Position Average for General Products") +
  xlab("Product Type") +
  ylab("Position")


```

>These boxplots compare positions for Amazon and partner products between general and specific search queries. The boxplots reveal that for general queries, Amazon typically has its own brand for general products and will rank it high; however, there are many partnering products that come up right behind it therefore the means are similar hovering around 1 and 3. As for specific searches, Amazon typically does not have its own brand, causing the average to dip around 19 whereas partner products average remains high around 2. This concludes that for general searches, Amazon tends to consequently position itself much higher; however, for specific searches Amazon typically lacks its own version and will position itself significantly lower.


### Question 2
- What is the distribution of positions for Amazon products compared to Non_Amazon products in search results?

We answer this by generating a graph of all Amazon products and the frequency of their position on the website. 
```{r, echo = F}
amazon %>%
    ggplot(aes(x=amazon)) +
    geom_bar(fill = 'blue', color = 'black') +
    ggtitle("Distribution of Positions for Amazon Products") +
    xlab("Position on Amazon") +
    ylab("Frequency") +
    geom_vline(xintercept = 3.5, color = 'red', linetype = 'dashed')

amazon %>%
    ggplot(aes(x=partner)) +
    geom_bar(fill = 'purple', color = 'black') +
    ggtitle("Distribution of Positions for Non_Amazon Products") +
    xlab("Position on Amazon") +
    ylab("Frequency") +
    scale_y_log10() #Added log scale; don't know if it makes it look better or worse
```

> Connecting to Q1, these two graphs show the contrast in product placement between Amazon and non_Amazon products. With its own products, Amazon tends to place its own products either at a premiere position (1st) or much lower (down to approx. 60th), whereas non_Amazon products are always placed in higher positions since third parties can produce more specialized products, also including general.

> Note that the tall bar ONLY accounts for situations in which Amazon's product is #1. While this may be reminicient of their product quality, it is not reflected in a general downward trend as the position increases. Instead, after position 1, the frequency is not consistent until a dropoff after 40.  This could suggest that Amazon does not place their items at #1 because they have the best quality, but rather that they want more viewer traffic over their products. 


### Question 3
- What is the difference in correlation for stars and reviews in products where Amazon is first compared to where non-Amazon products come first?

>We first filter products in which Amazon's position on the website is higher than the other two. Comparing stars to reviews, we generate a linear relationship between the two and graph the residuals based on our line of best fit. After that, we calculate the correlation coefficient relating our two variables. 

```{r, echo = F}
amazon_first = amazon %>% #select products in which Amazon comes first
  filter(amazon < partner) %>% 
  select(starts_with('amazon')) 

amazon_lm = lm(amazon_stars ~ amazon_reviews, data = amazon_first) 
#Correlation between stars and reviews

amazon_first = amazon_first %>% 
  add_residuals(amazon_lm)

ggplot(amazon_first) + #Residual plot with reviews on x, residuals on y
  aes(amazon_reviews, resid) + 
  geom_point(color = 'chartreuse4', size = 4) +
  scale_x_log10()+ 
  geom_hline(yintercept = 0, color = 'red', linetype = 'dashed') +
  xlab("Number of Reviews")+
  ylab("Residuals") +
  ggtitle("Residual Plot of Amazon Reviews", 
          subtitle = "Where Amazon had the First Product")

correlation1 = cor(amazon_first$amazon_stars, amazon_first$amazon_reviews)
correlation1
```

- The spread of the residuals in this plot is pretty evenly spread out, with a correlation coefficient of .27. While this is low, it's expected not to have a perfect correlation as most products are rated very close together. 

>Next, we do the same thing as before, but instead we look at products in which Amazon didn't come first. 

```{r, echo = F}
non_amazon_first = amazon %>% #Same as previous, but with non-amazon products (partner)
  filter(partner < amazon) %>% 
  select(starts_with('partner'))

non_amazon_lm = lm(partner_stars ~ partner_reviews, data = non_amazon_first)

non_amazon_first = non_amazon_first %>% 
  add_residuals(non_amazon_lm)

ggplot(non_amazon_first) + 
  aes(partner_reviews, resid) + 
  geom_point(color = 'orchid3', size = 4) +
  scale_x_log10() + 
  geom_hline(yintercept = 0, color = 'red', linetype = 'dashed') +
  xlab("Number of Reviews")+
  ylab("Residuals") +
  ggtitle("Residual Plot of Non-Amazon Reviews", 
          subtitle = "Where Non-Amazon Sellers had the First Product")

correlation2 = cor(non_amazon_first$partner_stars, non_amazon_first$partner_reviews)
correlation2
```

- The spread of the residuals in this plot is much worse than the previous one. Note how the dashed line 0 represents the middle of the spread, which is not shown in this graph. There is much more spread under 0 compared to the spread over 0. Also note that the correlation coefficient is 0.08, which is significantly lower than Amazon's 0.27.

####   Analysis 
>The correlation between number of stars and number of reviews is much stronger in products where Amazon was ranked higher than non-Amazon products. This suggests that there is a stronger relationship between number of reviews and quality of Amazon products compared to non-Amazon products. For the everyday consumer, this could mean that Amazon products are a more trustworthy brand, at the very least for the general products it offers.




## Methods
We conduct two hypothesis tests: the first one is that Amazon products are on average placed higher than third-party products, and the second is that Amazon partner products are on average placed higher than third-party products. With these hypothesis tests on the differences in means, we can present our numerical results as an answer to whether Amazon and partner products are placed higher than third-party products. We will then run summary statistics on stars and reviews comparing the same groups to see if our mean positions are justified, i.e. a product with a higher star rating and more reviews is placed higher than the rest. 

### Question 4
- Hypothesis Test: Are Amazon products on average placed higher than those from third-party sellers?

>We will answer this question by performing a series of mutations on our data in order to calculate two distinct values. First, we will find the difference in values between these two groups, and use these values to search for the true mean between them. In order to do that, we will construct a 95% confidence interval of where we think the true mean resides. Finally, we will calculate a t-statistic and use this to calculate a p-value to show how probable our hypothesized mean is compared to the true mean. For both tests, in order to test if one group is higher than the other, we will first need to test if both groups are similar. To test similarity, we will predict a mean of 0. If 0 is within the 95% confidence interval, we can assume that there is a probable chance it is the true mean. This is similar to our p-value. If the p-value is over 0.05, we can assume that there is a statistically significant chance that our predicted value could be the true mean. 

```{r, include = FALSE}
#Used pivot_longer to put name of retailer in its own column
amazon_longer = amazon %>% 
  select(amazon, partner, third_party) %>% 
  pivot_longer(c(amazon, partner, third_party), 
               names_to = 'retailer', values_to = 'position')
#summary of n, mean, and sd of each retailer
amazon_longer %>% 
  group_by(retailer) %>% 
  summarize(n = n(), mean = mean(position), sd = sd(position))
```

Model 
$$
X_i \ \sim \ F_1(\mu_1, \sigma_1), i = 1, ..., n_1 \\
Y_i \ \sim \ F_2(\mu_2, \sigma_2), i = 1, ..., n_2
$$
Hypothesis test
$$
H_0 : \mu_1 = \mu_2     \\
H_a : \mu_1 \ne \mu_2
$$
Our test statistic is $T$

If the null hypothesis is true, then:
$$ 
X_i = Y_i
$$

>After our calculations, we have constructed a confidence interval for the mean of our two groups:

```{r, echo=FALSE}
#Code modified from lizard examples in class to calculate 95% confidence interval of
#Amazon compared to third party
#Note that 0 is within the interval, which supports the null hypothesis
amazon_sum = amazon %>% 
  mutate(diff = amazon-third_party) %>% 
  summarize(n = n(),
            mean = mean(diff),
            sd = sd(diff),
            r = cor(amazon, third_party),
            t = mean/(sd/sqrt(n)))

ci_calc = amazon_sum %>% 
  mutate(se = sd/sqrt(n), 
         tmult = qt(0.975, n-1), 
         me = tmult*se,
         low = mean - me, 
         high = mean + me)

ci = ci_calc %>% 
  select(low, high)

ci
```

>We are 95% confident that the average difference in position between Amazon products and third party products is between -0.366 and 1.45. Note that this includes 0, which provides evidence supporting the null hypothesis that the two sets of products are similar. 

>Next, we calculate the p-value: 

```{r, echo=FALSE}
#Calculate two sided p-value
pvalue = 2*pt(-abs(amazon_sum$t), amazon_sum$n-1)
pvalue
```

>There is a good chance (p = 0.24) that the mean difference in position between Amazon products and third party products is 0. 

### Question 5
- Hypothesis Test: Are Amazon partner products on average placed higher than 
those from third-party sellers?

>We will then repeat our hypothesis test for comparing Third Party products vs Partner products.

Model 
$$
X_i \ \sim \ F_1(\mu_1, \sigma_1), i = 1, ..., n_1 \\
Y_i \ \sim \ F_2(\mu_2, \sigma_2), i = 1, ..., n_2
$$
Hypothesis test
$$
H_0 : \mu_1 = \mu_2     \\
H_a : \mu_1 \ne \mu_2
$$
Our test statistic is T

If the null hypothesis is true, then:
$$ 
X_i = Y_i
$$

>Same as last time, we now construct a confidence interval for the mean of our two groups. 

```{r, echo = FALSE}
#Repeated process with Third Party vs Partner
#Note that confidence interval is 14.0-15.2, which does not include 0
#This data rejects the null hypothesis
amazon_sum2 = amazon %>% 
  mutate(diff = third_party-partner) %>% 
  summarize(n = n(),
            mean = mean(diff),
            sd = sd(diff),
            r = cor(partner, third_party),
            t = mean/(sd/sqrt(n)))

ci_calc = amazon_sum2 %>% 
  mutate(se = sd/sqrt(n), 
         tmult = qt(0.975, n-1), 
         me = tmult*se,
         low = mean - me, 
         high = mean + me)

ci = ci_calc %>% 
  select(low, high)

ci
```
>We are 95% confident that the mean difference between third party products and partner products is between 14.0 and 15.2. Note that 0 is not within our confidence interval, as opposed to the previous hypothesis test. This, on the other hand, provides evidence contrary to the null hypothesis, which would then suggest that the two sets of products are not similar. 

>And again, we calculate our p value to test how probable our hypothesized mean of 0 is when compared to the mean of our dataset. 

```{r, echo=FALSE}
pvalue = 2*pt(-abs(amazon_sum2$t), amazon_sum2$n-1)
pvalue
```

>(Note that the calulated value was so small, our computer was forced to round to 0) Because of this, we are almost certain (p≈0) that the mean difference between third party products and partner products is not 0. 

## Analysis of Hypothesis Tests
According to our data, Amazon and its partners do not share any privileges of being placed higher on the website. There is strong evidence that they are equal in placement. Third-party products, however, do not share this privilege. They are on average placed way lower on the website than Amazon products, and there is very strong evidence to suggest that they are not equal. 

## Discussion
After running our methods on the dataset, we hope to achieve an answer as to how Amazon creates its search results, whether it places products from itself and its partners above products from third-party sellers. This is done after taking stars and review counts into consideration. The implications of this could be that Amazon manipulates its search results, where it favors third-party products the least despite having better stars and more reviews.

## Limitations of our Study
Especially with a project idea like studying Amazon search results, we encountered a lot of limitations in understanding the data. 

Originally we had planned to study whether Amazon places its own products higher on the the site, compared to non_Amazon products, but one limitation was that we didn't have search volume data. To retrieve the original data set from the github repo, would have been too big for this project span. Instead, using the subset of search results we had, we decided to change the study to examine how Amazon products are defined.

This data was collected in January 21st, 2021, although being recent, we do understand that company frameworks (including Amazon's) change over time, so search results could behave more differently in the present.

Amazon has had its fair share of criticism regarding fake reviews and star ratings, and many products included in our study could be subject to those actions, which misinterprets data and gives false information regarding the quality of the product. In a WIRED article we read on false reviews, it mentions that *"unscrupulous or desperate sellers pay for glowing reviews, and some offer gift cards or free products in exchange for five-star ratings. Occasionally, sellers will even solicit false negative reviews on competing products"*. This mass amount of false reviews and star ratings may not be an issue for more well known products, but for more specific search results, it can make a big difference in ranking.





## Conclusion



1. Does what you search matter in terms of product ranking? 

Based on Question 1 of our analysis, if your search is more general, then the results will be skewed towards products from Amazon and its partners; Amazon and partner products will be ranked near the top.Third-party products. On the other hand, if your search is more specific, then Amazon's products will be ranked significantly lower, but its partner products will stay consistently high.

2. Can we assume the type of search by simply looking at the product's ranking, stars and reviews on Amazon? 

Question 3 of our analysis told us that there was some correlation between stars and review count when Amazon products were first, while there was essentially no correlation between stars and review count when non-Amazon sellers were first. This inconsistency shows that  we can't reliably use ranking, stars, and review count to determine the type of search. 

This is further supported by Question 4 of our analysis, as our hypothesis testing comparing Amazon versus third-party products showed that the mean difference between their positions is 0. In other words, Amazon and third-party products are placed similarly. 

In addition, our hypothesis test in Question 5 of third-party versus partner products showed that we reject the null hypothesis. This means that the mean difference between their positions is not the same, telling us that third-party and partner products vary with where they are placed. The inconsistency between these two hypothesis tests tells us that positioning is inconsistent between our three types of sellers, so we can't reliably use product position to determine the type of search.

3. How do Amazon's products compare to those marketed by non-Amazon retailers?

Based on Question 2 of our analysis, which suggested that Amazon does not market their products higher based on quality, and Question 1 which suggested that Amazon markets their general products higher than specific ones, we had hypothesized that Amazon products were only placed higher to get more foot traffic and would not hold up in quality when compared to non-Amazon retailers. However, when we ran our hypothesis tests comparing the mean placement of Amazon vs Third Party, we found strong similarity between the two. This disproves our hypothesis, instead suggesting that there are other factors at play that determine the placement of Amazon products on their website. 

## Sources:
- https://medium.com/@Shoemaker703/how-does-amazon-select-the-top-product-in-its-search-results-and-why-is-it-usually-one-of-amazons-662d122591d0

- https://github.com/the-markup/investigation-amazon-brands

- https://landingcube.com/amazon-statistics/

- https://www.wired.com/story/how-to-spot-fake-reviews-amazon/#:~:text=The%20trouble%20is%20that%20many,negative%20reviews%20on%20competing%20products.
